Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence and Forensic Workflow

### 2009-03-11 paper/paper.tex
simson3 Ignore-this: 1c077e496d797e18ac61488f2b7b568a
Significant work done by Simson. Going to sleep now.

% Make sure we know its a draft for now
\usepackage{draftwatermark}
\SetWatermarkFontSize{3cm}
\usepackage{verbatim}
%\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage,linkcolor=black,citecolor=black,filecolor=black,urlcolor=black]{hyperref}
%\usepackage{natbib}
%\pagestyle{empty}
%\psdraft
%\baselineskip=20pt %Sets line spacing to 1 unit
%\bibliographystyle{elsart-num-names}
\bibliographystyle{IEEEtran}
\usepackage{cite}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1:}]}{\end{trivlist}}

@@

\subsection{AFF4 Terminology and New Concepts}

@@

\subsection{AFF4 Streams}
As discussed above, individual segments in an AFF4 evidence set can be
combined into a single logical object called a \emph{stream}. AFF4
supports an arbitrary number of streams per volume. The AFFLIB library
allows the segments in a stream to be operated on as if they were a
single file inside a file system by supporting the traditional
POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
\texttt{read()}.

@@

@@

\subsection{ZIP64}
For AFF4, we have decided to change AFF1's
original container file format to ZIP64\cite{zipspecs}. There are many
reasons for this decision:

\subsubsection{A Forensicly Sound Strategy For Updating ZIP64 Files}

\subsubsection{Segment merging}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Standard FIF segments}
The FIF specification defines a number of basic types, but extensions
are encouraged. Future versions of the specification might standardize
further types. Implementations MUST implement the basic types.

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image. A TCP/IP stream is simply a collection of
payloads from selected network packets. Sometimes the same data may be
viewed in a number of ways - for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation. It is, therefore, beneficial to define a way in which
the underlying storage system can apply transformations to existing
data to produce transformed data.

For example, Zero Storage Carving \cite{Meijer2006}, is a way of
specifying carved files in terms of a sequence of blocks taken from
the image. \cite{Cohen2007} extended this concept to an arbitrary
mapping function. This is the approach chosen in this specification
due to its flexibility.

@@

Denoting the stream offset by $x$, and the target offset by $y$, the
map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
byte at a mapped stream offset $x$ can then be satisfied by reading a
byte from target $T_i$ at offset $y$ given by:
\begin{eqnarray}
y = (x - X_i) + Y_i & &
\forall x \in \left [X_i, X_i+1 \right )
\end{eqnarray}

@@

Where $mod$ is the modulus function, while $floor$ signifies integer
division. For example consider the following, which corresponds to a 3
disk RAID-5 array:
\begin{verbatim*}
stream_period=393216
target_period=196608

0,0,1
65536,0,0
131072,65536,2
196608,65536,1
262144,131072,0
327680,131072,2
\end{verbatim*}

\subsection{The Overlay stream}
The Overlay stream is a variant of the {\em Image} stream in that it
allows for images to be stored in (possibly compressed) chunks. The
main utility of the Overlay stream is in providing transparent access
to other forensic formats, such as EWF or AFF. These formats also use
chunk based compression to provide the ability to randomly seek while
reading. To access these file formats we only need to have an index of
each chunk offset, and then directly use the overlayed files. The {\em
ewf2fif} tool provides such an overlay feature for EWF volumes by
employing libewf to scan the internal data structures.

Therefore, the {\em chunk\_size} attribute must be specified. In
addition the {\em compression\_type} attribute MAY be specified. The
attribute should contain an integer corresponding to a suitable Zip
compression scheme (e.g. DEFLATE). If not specified it is assumed
that no compression is used.

@@

It is possible to refer to the current file by providing a target of
``file://.''. This can be useful for modifying an existing EWF file to
be a FIF file. Since Zip files are normally read from the end of the
file, but EWF and AFF are both read from the front of the file it may
be possible to append a FIF file to the end of an EWF file without
interfering with the overlayed file. In this case the converted file
can still be used as an EWF file without change, but it can also be
used as a FIF file.

@@

In the above figure a segment is written to the FIF volume which
contains a number of chunks back to back. An overlay stream then uses
direct references to each compressed chunk within the coalesced
segment. The target is then specified as the segment name. Although
this technique is valid, it is generally discouraged since it breaks
the appealing simplicity of the standard Image Stream.  It is no
longer possible to simply unzip all the chunks and concatenate them
together to recover the original image.

\subsection{The Private stream}
Often forensic software needs to store internal state, such as the
current state of the GUI, or internal data structures which may be
required for caching. Usually, software create additional cache files
to maintain this information. It is advantageous to be able to store
these in the evidence file itself. The evidence file can then be
re-opened by the software at a later stage and private data can be
retrieved and used directly. Such private application data can be
stored in a Private stream. There is no specification of what can be
stored in the private stream, and applications with private streams
should be able to store arbitrary segments in any format at all.

\subsection{Encrypted Streams}
Encryption is an important property in an evidence file format
\cite{AFF}. In particular with the FIF format, multiple streams may be
present in the file set, and often different access levels are
desired. For example, for evidence set containing both network
captures and disk images it may be desirable to limit access to
streams based on legal authorizations.

@@

The Encrypted Stream ({\em type=encrypted}) SHOULD be implemented by
FIF implementations, but may be omitted with an obvious loss of
functionality (Encryption may be prohibited or restricted in some
jurisdictions).

The Encrypted Stream is similar to the Image Stream above in that data
is written out in chunks (although its probably not effective to
compress these chunks at all). Each chunk is to be encrypted upon
writing and transparently decrypted using a scheme declared in the
{\em scheme} attribute.

The Encrypted Stream may contain any data at all. It is useful
however, to store an actual FIF volume within the Encrypted
stream. This provides block level encryption for the contained FIF
volume. It is recommended that Encrypted Streams containing a FIF
volume set an attribute ({\em content-type=application/x-fif-volume})
to allow implementations to automatically recognize volumes. This
approach is illustrated in Figure \ref{crypted_fif}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
  \caption{Embedding an encrypted FIF volume within an Encrypted Stream.}
  \label{crypted_fif}
  \end{center}
\end{figure}

The result is that a number of FIF volume are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the Container Volume may contain several Encrypted
Streams and therefore contain multiple Encrypted FIF
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted scheme. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be merged.

Implementations should create embedded FIF volumes as extensions to
the container volumes (i.e. with the same UUID). Thus when embedded
volumes are opened, their streams become automatically viewable as
part of the evidence set.

@@

Implementations MUST specify the salt as a stream attribute named {\em
salt} in base 64 encoding. Implementations are free to specify any API
for passing the Pre-Shared Key, for example the PSK can be specified
in a configuration file, typed in at the terminal, passed in using
process environment variables, or encoded within the FIF filename
itself.

\subsection{Optional features}
Implementations may implement the following features to simplify the
interfaces.

\subsubsection{Automated volume loading}
FIF volumes are effectively independent, as each volume does not need
to know anything about other volumes. This makes it necessary to
specify in advance all volumes for the implementation to merge them
all. In particular it may be necessary to specify encrypted volumes
(which may not be read if the user has no authorization for the
Encrypted Stream).

As an optional enhancement FIF specifies a volume attribute named {\em
volume} (i.e. an attribute present in the top level ``properties''
segment). This is a target reference (see section
\ref{target_reference}) for other volumes. 
Implementations should attempt to open external references (e.g. using
the {\em file://} convention) before internal references. As new
volumes are merged their volume attributes must be scanned for
additional references. Each volume must list all other known volumes
as volume attributes.

In this way, encrypted volumes, and external volumes can be located
and found automatically.

\section{Implementation Recommendations}
Since the individual chunks are simply files in a Zip archive, any zip
library can be utilized for reading these chunks. However, this may
not be efficient since simple Zip implementations might rescan the
Central Directory for each segment request. Since in practice there
can be a large number of chunks in each volume, it is recommended that
implementations build an index of offsets and length of each chunk
into the file. This kind of index is very similar to the strategy
other file formats use \cite{EWF}. The index may then be saved to the
file by use of a Private stream.

Zip64 specifies duplicate data in the Central Directory entry for each
segment and its File Header (such as filename, size, CRC etc). Many
implementations only populate this information in one of these
places. In the interest of robustness, it is recommended that FIF
implementations populate both places at the slight expense of higher
overheads, because this can be used to recover and correct damaged
files. If the central directory is lost, it is possible to scan
through the zip file and repair the central directory from the File
headers.

The underlying Zip implementation should support multiple writers
simultaneously by interleaving their writes to the file. For example,
for two streams being written concurrently, the new segments can be
added to the same archive in any order. In addition, the volume may be
closed and a new volume opened without writers being aware of
this. Since all segments from all volumes are merged, streams can be
split arbitrarily across volumes in any way. It is recommended that
stream properties be written to each volume to ensure at least some of
the stream can be used even if some volumes are lost.

@@

\section{Conclusion and Future Work}
\subsection{Expanding AFFLIB for AFF4}
Although we do not specify the exact API, implementations
MUST make a stream opened for reading have at a minimum a read, seek
and tell methods. Seeking MUST support the whence argument as
described in the standard C implementation (i.e. allow seeking to the
end of the stream).

@@

>>>>

% Make sure we know its a draft for now
\usepackage{draftwatermark}
\SetWatermarkFontSize{3cm}
\usepackage{fancyvrb}
%\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage,linkcolor=black,citecolor=black,filecolor=black,urlcolor=black]{hyperref}
%\usepackage{natbib}
%\pagestyle{empty}
%\psdraft
%\baselineskip=20pt %Sets line spacing to 1 unit
%\bibliographystyle{elsart-num-names}
\bibliographystyle{IEEEtran}
\usepackage{cite}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1:}]}{\end{trivlist}}

@@

\subsection{New Concepts and Terminology}

@@

\subsection{Streams}
As discussed above, individual segments in an AFF4 evidence set can be
combined into a single logical object called a \emph{stream}. AFF4
supports an arbitrary number of streams per volume. The AFFLIB library
allows the segments in a stream to be operated on as if they were a
single file inside a file system by supporting the traditional
POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
\texttt{read()}.

@@

\subsection{ZIP64}
For AFF4, we have changed AFF1's original container file format to
ZIP64\cite{zipspecs}. There are many reasons for this decision:

\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}

\section{Standard AFF4 Streams}
In this section we discuss specific stream types that we have created
in our initial AFF4 implementation.

@@

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image, while a TCP/IP stream is simply a collection of
payloads from selected network packets. 
Sometimes the same data may be
viewed in a number of ways---for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation.  Zero Storage Carving \cite{Meijer2006} is a way of
specifying carved files in terms of a sequence of blocks taken from
the image; Cohen extended this concept to an arbitrary
mapping function\cite{Cohen2007} .

AFF4 defines a common way in
which the underlying storage system can apply transformations to
existing data to produce transformed data. This transformation is
performed using a \emph{map stream}.

@@

Denoting the stream offset by $x$, and the target offset by $y$, the
Map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
byte at a mapped stream offset $x$ can then be satisfied by reading a
byte from target $T_i$ at offset $y$ given by:
\begin{eqnarray}
y = (x - X_i) + Y_i & &
\forall x \in \left [X_i, X_i+1 \right )
\end{eqnarray}

@@

Where $mod$ is the modulus function and $floor$ signifies integer
division. For example consider Figure~\ref{map}, which corresponds to a 3
disk RAID-5 array.

\begin{figure}
\begin{Verbatim}[frame=single]
stream_period=393216
target_period=196608

0,0,1
65536,0,0
131072,65536,2
196608,65536,1
262144,131072,0
327680,131072,2
\end{Verbatim}
\caption{A Map stream that corresponds to a 3 disk RAID-5 array}\label{ref}
\end{figure}

\subsection{The Overlay Stream}
The Overlay Stream is a variant of the {\em Image} stream that
provides a layer of indirection between requests made for specific
bytes (or chunks or bevys) and the actual location where the requested
information is stored. The main utility of the Overlay stream is in
providing transparent access to legacy forensic formats, such as EWF.

To access these legacy file formats we only need to have an index of
each chunk offset, and then directly use the overlayed files.  The {\em
ewf2aff} tool provides such an overlay feature for EWF volumes by
employing libewf to scan the internal data structures.

Overlay stream require that a \texttt{chunk\_size} attribute be specified. In
addition the {\em compression\_type} attribute may be specified. The
attribute should contain an integer corresponding to a suitable Zip
compression scheme (e.g. DEFLATE). If not specified it is assumed
that no compression is used.

@@

It is possible to refer to the current file by providing a target of
``file://.''. This can be useful for modifying an existing EWF file to
be a AFF4 file. Since Zip files are normally read from the end of the
file, but EWF are both read from the front of the file, it may
be possible to append a AFF4 Central Directory to the end of an EWF file without
interfering with the overlayed file. In this case the converted file
can still be used as an EWF file without change, but could also be
used as an AFF4 file.

@@

In the above figure a segment is written to the AFF4 volume which
contains a number of chunks back to back. An overlay stream then uses
direct references to each compressed chunk within the coalesced
segment. The target is then specified as the segment name. Although
this technique is valid, it is generally discouraged since it breaks
the appealing simplicity of the standard Image Stream.  It is no
longer possible to simply unzip all the chunks and concatenate them
together to recover the original image.

\subsection{Encrypted Streams}
Encryption is an important property in an evidence file format. In
particular, multiple streams may be present in the file set, and often
different access levels are desired. For example, for evidence set
containing both network captures and disk images it may be desirable
to limit access to streams based on legal authorizations.

@@

AFF4 therefore introduces a new encryption scheme, the Encrypted
Stream.  The Encrypted Stream is a stream that encapsulates other
streams on a bevy-by-bevy basis. Each bevy transparently encrypted upon
writing and decrypted on reading. The choice of algorithm and mode is declared in the
Stream's \texttt{scheme} attribute.

The Encrypted Stream may contain any data at all. It is useful
however, to store an actual AFF4 volume within the Encrypted
stream. This provides block level encryption for the contained AFF4
volume. It is recommended that Encrypted Streams containing a AFF
volume set an attribute (\texttt{content-type=application/x-aff-volume})
to allow implementations to automatically recognize volumes. This
approach is illustrated in Figure~\ref{crypted_fif}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
  \caption{Embedding an encrypted AFF4 volume within an Encrypted Stream.}
  \label{crypted_fif}
  \end{center}
\end{figure}

The result is that a number of AFF4 volumes are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the outer Volume may contain several Encrypted
Streams and therefore contain multiple AFF4 Encrypted 
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted scheme. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be merged.

@@

Implementations MUST specify the salt as a stream attribute named {\em
salt} in base 64 encoding. Implementations are free to specify any API
for passing the Pre-Shared Key, for example the PSK can be specified
in a configuration file, typed in at the terminal, passed in using
process environment variables, or encoded within the AFF4 filename
itself.

\subsection{Private Streams}
Often forensic software needs to store internal state, such as the
current state of the GUI, or internal data structures which may be
required for caching. Usually, software create additional cache files
to maintain this information. It is advantageous to be able to store
these in the evidence file itself. The evidence file can then be
re-opened by the software at a later stage and private data can be
retrieved and used directly. Such private application data can be
stored in a Private stream. There is no specification of what can be
stored in the private stream, and applications with private streams
should be able to store arbitrary segments in any format at all.

\section{Implementing AFF4 in AFFLIB}
Although there are numerous ZIP implementations available
today, we have created our own implementation using a combination of
available open source technology and our own novel contributions. We
are in the process of integrating this code into the AFFLIB library
and performing extensive regression testing.

There are many reasons to develop our own ZIP64 implementation for
AFF4:

\begin{itemize}
\item Most ZIP implementations do not implement the ZIP64
  extensions. These extensions are required to support Evidence
  Volumes larger than 2GB.
\item Simple Zip implementations might rescan the
Central Directory for each segment request. Since in practice there
can be a large number of segments in a volume, it is advisable to have
a ZIP64 implementation that is optimized to storing thousands (or even
hundreds of thousands) of segments in an efficient data structure. 
\item While the ZIP specification duplicates data found in the Central
  Directory entry in each File Header (such as filename, size, CRC
  etc), many implementations that we have examined only populate this
  information in one of these places. In the interest of robustness,
  we wanted to assure that data stored in both locations would be
  populated to allow recovery of at least \emph{some} evidence that
  might exist in damaged volumes. If central directory is lost, it is
  possible to scan through the zip file and repair the central
  directory from the File headers.

\item Our implementation supports simultaneous access by multiple
  readers and writers. Our system allows concurrency by locking the
  file and interleaving write operations. Managing simultaneous
  writers is simplified through our update strategy which only writes
  new segments to the file, never modifying old segments. 
\item AFF4 ``properties'' have specific semantics above and beyond
  normal segments. For example, AFF4 calls for all of the property
  segments associated with a file to be merged together. Furthermore,
  when updating the properties in a multi-volume Set, it is useful to
  write the same properties to each volume to ensure at least some of
the stream can be used even if some volumes are lost. Such
functionality is not present in standard ZIP64 implementations.
\end{itemize}

@@

\section{Conclusion and Future Work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
